animbook: Visualizing Changes in Performance Measures and Demographic Affiliations using Animation

### 2025-04-30 animbook-journal.Rmd
cleaning and glimpse function give latex error

```{r tbl-scaling, eval=knitr::is_html_output()}
scaling_table |> 
  knitr::kable(caption = "A summary of different scaling methods applied to the dataset.")
```

```{r tbl-scaling-la, eval=knitr::is_latex_output()}
scaling_table |> 
  knitr::kable(caption = "A summary of different scaling methods applied to the dataset.") |> 
  kableExtra::column_spec(8, width = "30em")
```

@@

```{r, echo=TRUE}
a <- anim_prep_cat(cat_change, 
                   id = id, 
                   values = qnt, 
                   time = time, 
                   group = gp)
class(a)
glimpse(a)
```

@@

p <- wallaby_plot(accounting,
                  subset = "bottom",
                  relation = "many_one",
                  height = 1,
                  size = 2,
                  width = 100,
                  total_point = 1000)

@@

p2 <- anim_animate(p)
```

```{r, eval=FALSE}
gganimate::anim_save("figures/exit.gif", p2, height = 8, fps = 30,
                     nframes = 400, width = 9, units = "in", res = 150)

@@

>>>>

```{r tbl-scaling}
scaling_table |> 
  knitr::kable(caption = "A summary of different scaling methods applied to the dataset.")
```

@@

```{r, echo=TRUE}
a <- anim_prep_cat(cat_change, 
                   id = id, 
                   values = qnt, 
                   time = time, 
                   group = gp)
# class(a)
# glimpse(a)
```

@@

wal_p <- wallaby_plot(accounting,
                  subset = "bottom",
                  relation = "many_one",
                  height = 1,
                  size = 2,
                  width = 100,
                  total_point = 1000)

@@

wal_p2 <- anim_animate(wal_p)
```

```{r, eval=FALSE}
gganimate::anim_save("figures/exit.gif", wal_p2, height = 8, fps = 30,
                     nframes = 400, width = 9, units = "in", res = 150)

@@

gganimate::animate(wal_p2)
```

### 2025-04-30 animbook-journal.tex
cleaning and glimpse function give latex error

\section{Introduction}\label{intro}

@@

\section{Explanation of the New York Times visualization}\label{NYTvis}

@@

\section{Data}\label{data}

@@

\section{Animation tools}\label{animation}

@@

\section{Visualization design}\label{design}

@@

\section{Software}\label{software}

\subsection{Installation}\label{installation}

@@

\subsection{Overview of functions}\label{overview-of-functions}

@@

\subsubsection{Data pre-processing}\label{data-pre-processing}

@@

\caption{\label{tab:tbl-scaling-la}A summary of different scaling methods applied to the dataset.}
\centering
\begin{tabular}[t]{l|>{\raggedleft\arraybackslash}p{30em}|l|r|r|r|r|r}
\hline
id & time & gp & values & rank & rank\_group & absolute & absolute\_group\\
\hline
1 & 2020 & X & 4254.6677 & 3 & 3 & 3 & 3\\
\hline
1 & 2023 & X & 3357.0312 & 2 & 2 & 4 & 4\\
\hline
14 & 2020 & X & 6763.4772 & 2 & 2 & 2 & 2\\
\hline
14 & 2023 & X & 1197.4183 & 4 & 5 & 5 & 5\\
\hline
100 & 2020 & X & 6864.4950 & 2 & 2 & 2 & 2\\
\hline
100 & 2023 & X & 2321.3183 & 3 & 3 & 4 & 4\\
\hline
21 & 2020 & Y & 698.2479 & 5 & 5 & 5 & 5\\
\hline
21 & 2023 & Y & 3970.0503 & 2 & 2 & 4 & 3\\
\hline
106 & 2020 & Y & 4109.6530 & 3 & 3 & 3 & 3\\
\hline
106 & 2023 & Y & 2865.6471 & 3 & 3 & 4 & 4\\
\hline
148 & 2020 & Y & 7173.7712 & 2 & 2 & 2 & 2\\
\hline
148 & 2023 & Y & 4217.4130 & 1 & 1 & 3 & 3\\
\hline
\end{tabular}
\end{table}

@@

These are only the initial steps in formatting the data into a category. Now that there is a method to transform the data from the raw into a categorized format, the next step is to modify it into an animated data structure. It is carried out by assigning a frame to each individual observation, ensuring that each ID does not contain repeated frame values. It lets the \CRANpkg{gganimate} or \CRANpkg{plotly} perceive where the observation would be on the plot at a given frame, as seen in Figure \ref{fig:animated-diagram}.

@@

\caption{\label{tab:tbl-cat-la}The argumenst for the anim\_prep\_cat function.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{30em}}
\hline
Argument & Description\\
\hline
data & The numerical data to be prepared for visualization.\\
\hline
id & The column name in the data that will be used to identify an individual over time. For example, company name.\\
\hline
values & The column name in the data that will be used for the y-axis (must be numerical).\\
\hline
time & The column name in the data represents the time variable (x-axis).\\
\hline
group & The column name in the data for distinguishes between the values group (for example, country). In the visualization, this will be used as a color argument.\\
\hline
ncat & The number of categories to create for scaling values. The default for this function is 5 categories.\\
\hline
breaks & A vector of breaks for creating bins. If this is not provided, the bins will have equal size.\\
\hline
label & A vector of labels to be used for the y-axis in the visualization. If this is not provided, the labels will be the position on the y-axis.\\
\hline
group\_scaling & The column name in the data that will be used for grouping. Allow the function to isolate the calculation between the groups.\\
\hline
scaling & The scaling method to be used; "rank" or "absolute". The default scaling is "rank".\\
\hline
\end{tabular}
\end{table}

The arguments for the \texttt{anim\_prep()} and \texttt{anim\_prep\_cat()} functions are listed in Tables \ref{tab:tbl-num-la} and \ref{tab:tbl-cat-la} respectively. The \texttt{data}, \texttt{id}, \texttt{values}, and \texttt{time} arguments are required for both functions.

The \texttt{ncat}, \texttt{breaks}, \texttt{group\_scaling}, and \texttt{scaling} arguments are for the user to customize the data scaling calculation in the \texttt{anim\_prep()} function. In the \texttt{anim\_prep\_cat} function, the users can customize the order of the category to be shown on the y-axis using the \texttt{order} argument.

Then, for further customization regarding how the final visualization looks. The \texttt{group}, \texttt{label}, and \texttt{time\_dependent} arguments can be adjusted.

@@

\subsubsection{Plotting function}\label{plotting-function}

@@

\caption{\label{tab:unnamed-chunk-8}The arguments for the wallaby\_plot function.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{30em}}
\hline
Argument & Description\\
\hline
data & The categorized data returned from the prep function.\\
\hline
group\_palette & The vector of the palette used by the function to supply the color to each group.\\
\hline
shade\_palette & The vector of the palette used by the function to supply the color to the shaded area.\\
\hline
rendering & The choice of method used to create and display the plot, either gganimate or plotly. Default is gganimate.\\
\hline
time\_dependent & Should the observations entered the visualization at the same time. Default is FALSE\\
\hline
subset & A character string specifying the variable used for subsetting the data. The "top" and "bottom" strings can also be used in this argument. Default is "top".\\
\hline
relation & The choice of relationship for the values to display on the plot, either "one\_many." or "many\_one". Default is "one\_many".\\
\hline
total\_point & The number of points the users want for the wallaby plot. Default is NULL, which is the number of points equal to the original.\\
\hline
height & The proportion of the area occupied by the observations in the shaded areas. Default is 0.6.\\
\hline
width & The distance between the first and last observation in the animation. Default is 50.\\
\hline
size & The point size. Default is 2.\\
\hline
alpha & The opacity of the proportional shaded areas. Default is 0.1.\\
\hline
\end{tabular}
\end{table}

\subsubsection{Animating function}\label{animating-function}

@@

\subsection{Example usage}\label{example-usage}

The \texttt{cat\_change} data in the package is used to illustrate how to use this package. It has five categories, A to E, over two time points. Since the data is already in the categorized data structure, the \texttt{anim\_prep\_cat()} function is used. The output from this function is a class called \texttt{categorized}.

\begin{verbatim}
a <- anim_prep_cat(cat_change, 
                   id = id, 
                   values = qnt, 
                   time = time, 
                   group = gp)
class(a)
\end{verbatim}

\begin{verbatim}
#> [1] "tbl_df"      "tbl"         "data.frame"  "categorized"
\end{verbatim}

\begin{verbatim}
glimpse(a)
\end{verbatim}

\begin{verbatim}
#> Rows: 400
#> Columns: 5
#> $ id    <fct> 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 1~
#> $ time  <int> 2020, 2020, 2020, 2020, 2020, 2020, 2020, 2020, 2020, 2020, 2020~
#> $ qtile <dbl> 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5~
#> $ label <chr> "A", "A", "A", "A", "A", "A", "A", "A", "A", "A", "A", "A", "A",~
#> $ group <fct> X, X, X, X, X, X, X, X, X, X, X, X, X, X, X, X, X, X, X, X, Y, Y~
\end{verbatim}

@@

Figure \ref{fig:top-many}, shows the flexibility to use various options, such as \texttt{subset} or \texttt{relation}, to customize the plot appearance. Here \texttt{subset\ =\ "top"} is used to specify category A for the final time point, and \texttt{relation\ =\ "many\_one"} is used to specify that all categories flowing into A at time point 2 are shown.

@@

\section{Applications}\label{applications}

@@

\subsection{Accounting database: osiris}\label{accounting-database-osiris}

@@

p2 <- anim_animate(p)
\end{verbatim}

@@

@@

\subsection{Voter behavior}\label{voter-behavior}

@@

\section{Discussion}\label{discussion}

@@

\section*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

>>>>

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

@@

\hypertarget{NYTvis}{%
\section{Explanation of the New York Times visualization}\label{NYTvis}}

@@

\hypertarget{data}{%
\section{Data}\label{data}}

@@

\hypertarget{animation}{%
\section{Animation tools}\label{animation}}

@@

\hypertarget{design}{%
\section{Visualization design}\label{design}}

@@

\hypertarget{software}{%
\section{Software}\label{software}}

\hypertarget{installation}{%
\subsection{Installation}\label{installation}}

@@

\hypertarget{overview-of-functions}{%
\subsection{Overview of functions}\label{overview-of-functions}}

@@

\hypertarget{data-pre-processing}{%
\subsubsection{Data pre-processing}\label{data-pre-processing}}

@@

\caption{\label{tab:tbl-scaling}A summary of different scaling methods applied to the dataset.}
\centering
\begin{tabular}[t]{l|r|l|r|r|r|r|r}
\hline
id & time & gp & values & rank & rank\_group & absolute & absolute\_group\\
\hline
1 & 2020 & X & 4254.6677 & 3 & 3 & 3 & 3\\
\hline
1 & 2023 & X & 3357.0312 & 2 & 2 & 4 & 4\\
\hline
14 & 2020 & X & 6763.4772 & 2 & 2 & 2 & 2\\
\hline
14 & 2023 & X & 1197.4183 & 4 & 5 & 5 & 5\\
\hline
100 & 2020 & X & 6864.4950 & 2 & 2 & 2 & 2\\
\hline
100 & 2023 & X & 2321.3183 & 3 & 3 & 4 & 4\\
\hline
21 & 2020 & Y & 698.2479 & 5 & 5 & 5 & 5\\
\hline
21 & 2023 & Y & 3970.0503 & 2 & 2 & 4 & 3\\
\hline
106 & 2020 & Y & 4109.6530 & 3 & 3 & 3 & 3\\
\hline
106 & 2023 & Y & 2865.6471 & 3 & 3 & 4 & 4\\
\hline
148 & 2020 & Y & 7173.7712 & 2 & 2 & 2 & 2\\
\hline
148 & 2023 & Y & 4217.4130 & 1 & 1 & 3 & 3\\
\hline
\end{tabular}
\end{table}

@@

These are only the initial steps in formatting the data into a category. Now that there is a method to transform the data from the raw format into a categorized format, the next step is to modify it into an animated data structure. It is carried out by assigning a frame to each individual observation, ensuring that each ID does not contain repeated frame values. It lets the \CRANpkg{gganimate} or \CRANpkg{plotly} perceive where the observation would be on the plot at a given frame, as seen in Figure \ref{fig:animated-diagram}.

@@

\caption{\label{tab:tbl-cat-la}The arguments for the anim\_prep\_cat function.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{30em}}
\hline
Argument & Description\\
\hline
data & The numerical data to be prepared for visualization.\\
\hline
id & The column name in the data that will be used to identify an individual over time. For example, company name.\\
\hline
values & The column name in the data that will be used for the y-axis (must be numerical).\\
\hline
time & The column name in the data represents the time variable (x-axis).\\
\hline
group & The column name in the data for distinguishes between the values group (for example, country). In the visualization, this will be used as a color argument.\\
\hline
ncat & The number of categories to create for scaling values. The default for this function is 5 categories.\\
\hline
breaks & A vector of breaks for creating bins. If this is not provided, the bins will have equal size.\\
\hline
label & A vector of labels to be used for the y-axis in the visualization. If this is not provided, the labels will be the position on the y-axis.\\
\hline
group\_scaling & The column name in the data that will be used for grouping. Allow the function to isolate the calculation between the groups.\\
\hline
scaling & The scaling method to be used; "rank" or "absolute". The default scaling is "rank".\\
\hline
\end{tabular}
\end{table}

The arguments for the \texttt{anim\_prep()} and \texttt{anim\_prep\_cat()} functions are listed in Tables \ref{tab:tbl-num-la} and \ref{tab:tbl-cat-la} respectively. The \texttt{data}, \texttt{id}, \texttt{values}, and \texttt{time} arguments are required for both functions. The \texttt{ncat}, \texttt{breaks}, \texttt{group\_scaling}, and \texttt{scaling} arguments are for the user to customize the data scaling calculation in the \texttt{anim\_prep()} function. In the \texttt{anim\_prep\_cat} function, the users can customize the order of the category to be shown on the y-axis using the \texttt{order} argument. Then, for further customization regarding how the final visualization looks. The \texttt{group}, \texttt{label}, and \texttt{time\_dependent} arguments can be adjusted.

@@

\hypertarget{plotting-function}{%
\subsubsection{Plotting function}\label{plotting-function}}

@@

\caption{\label{tab:unnamed-chunk-9}The arguments for the wallaby\_plot function.}
\centering
\begin{tabular}[t]{l|>{\raggedright\arraybackslash}p{30em}}
\hline
Argument & Description\\
\hline
data & The categorized data returned from the prep function.\\
\hline
group\_palette & The vector of the palette used by the function to supply the color to each group.\\
\hline
shade\_palette & The vector of the palette used by the function to supply the color to the shaded area.\\
\hline
rendering & The choice of method used to create and display the plot, either gganimate or plotly. Default is gganimate.\\
\hline
time\_dependent & Should the observations entered the visualization at the same time. Default is FALSE\\
\hline
subset & A character string specifying the variable used for subsetting the data. The "top" and "bottom" strings can also be used in this argument. Default is "top".\\
\hline
relation & The choice of relationship for the values to display on the plot, either "one\_many." or "many\_one". Default is "one\_many".\\
\hline
total\_point & The number of points the users want for the wallaby plot. Default is NULL, which is the number of points equal to the original.\\
\hline
height & The proportion of the area occupied by the observations in the shaded areas. Default is 0.6.\\
\hline
width & The distance between the first and last observation in the animation. Default is 50.\\
\hline
size & The point size. Default is 2.\\
\hline
alpha & The opacity of the proportional shaded areas. Default is 0.1.\\
\hline
\end{tabular}
\end{table}

\hypertarget{animating-function}{%
\subsubsection{Animating function}\label{animating-function}}

@@

\hypertarget{example-usage}{%
\subsection{Example usage}\label{example-usage}}

The \texttt{cat\_change} data in the \CRANpkg{animbook} package is used to illustrate how to use this package. It has five categories, A to E, over two time points. Since the data is already in the categorized data structure, the \texttt{anim\_prep\_cat()} function is used. The output from this function is a class called \texttt{categorized}.

\begin{verbatim}
a <- anim_prep_cat(cat_change, 
                   id = id, 
                   values = qnt, 
                   time = time, 
                   group = gp)
# class(a)
# glimpse(a)
\end{verbatim}

@@

Figure \ref{fig:top-many} shows the flexibility to use various options, such as \texttt{subset} or \texttt{relation}, to customize the plot appearance. Here \texttt{subset\ =\ "top"} is used to specify category A for the final time point, and \texttt{relation\ =\ "many\_one"} is used to specify that all categories flowing into A at time point 2 are shown.

@@

\hypertarget{applications}{%
\section{Applications}\label{applications}}

@@

\hypertarget{accounting-database-osiris}{%
\subsection{Accounting database: osiris}\label{accounting-database-osiris}}

@@

wal_p2 <- anim_animate(wal_p)
\end{verbatim}

@@

gganimate::animate(wal_p2)
\end{verbatim}

@@

\hypertarget{voter-behavior}{%
\subsection{Voter behavior}\label{voter-behavior}}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}
